---
title: "01 Explore temperature data in rivers"
output: html_document
---

# Get the number of temperature 

```{r}
library(dplyr)
library(niRvana)
library(purrr)
# install.packages("ODBC")
# library("odbc")
# help(package = "odbc")
```


```{r}
set_credentials()
```

## 1. Starting from the parameter/method end  
* Kind of blind end, as there are quite many WATER_CHEMISTRY_VALUES with temperature (5.5 million)    
* But we get 'df_method$METHOD_ID', which is the 73 IDs used for temperature. We will use this later (in part 2d)  
```{r}
df_param_def <- get_nivabase_selection("PARAMETER_ID, NAME", "WC_PARAMETER_DEFINITIONS", 
                                       "NAME", "Temperatur", values_are_text = TRUE)
nrow(df_param_def)

df_param_meth <- get_nivabase_selection("PARAMETER_ID, METHOD_ID", "WC_PARAMETERS_METHODS", 
                                       "PARAMETER_ID", df_param_def$PARAMETER_ID)
nrow(df_param_meth)

df_method <- get_nivabase_selection("METHOD_ID, NAME, UNIT", "METHOD_DEFINITIONS", 
                                  "METHOD_ID", df_param_meth$METHOD_ID)
nrow(df_method)

# VERY slow (I stopped it after 25 minutes)
# df_values <- get_nivabase_selection("METHOD_ID, WATER_SAMPLE_ID", "WATER_CHEMISTRY_VALUES", 
#                                    "METHOD_ID", df_param_meth$METHOD_ID)
# nrow(df_values)

get_nivabase_selection("count(*)", "WATER_CHEMISTRY_VALUES", 
                       "METHOD_ID", df_param_meth$METHOD_ID)


```

## 2. Starting from the station end
### a. Get river stations
```{r}
df_stationtype <- get_nivabase_selection("STATION_TYPE_ID, STATION_TYPE", "STATION_TYPES",
                                         "STATION_TYPE", "Elv", values_are_text = TRUE)
nrow(df_stationtype) # 1

df_stations <- get_nivabase_selection("STATION_TYPE_ID, STATION_ID", "STATIONS",
                                         "STATION_TYPE_ID", df_stationtype$STATION_TYPE_ID)
nrow(df_stations)  # 4388
head(df_stations)
```

### b. Test how to count using SQL   
This is just testing
Code adapted from https://stackoverflow.com/a/20120049/1734247  
```{r}
# test: get number of water samples, given station ID
#   stations (left join) samples
#   select by STATION_ID in stations
#   output: 1 line per STATION_ID
qry <- c("select stations.STATION_ID, count(samples.WATER_SAMPLE_ID) as count_samples",
         "from NIVADATABASE.STATIONS stations left join NIVADATABASE.WATER_SAMPLES samples" ,
         "on stations.STATION_ID = samples.STATION_ID ",
         "where stations.STATION_ID in (7579,7580,7581,7581) ",
         "group by stations.STATION_ID") %>% 
  paste(collapse = " ")
# edit(qry)
get_nivabase_data(qry)


# test: get number of chemistry values, given station ID and METHOD_ID
#   stations (left join) samples (left-join) chemical values
#   select by STATION_ID in stations + METHOD_ID in the chemical values
#   output: 1 line per STATION_ID
qry <- c("select stations.STATION_ID, count(chemval.VALUE_ID) as count_chem",
         "from (NIVADATABASE.STATIONS stations left join NIVADATABASE.WATER_SAMPLES samples" ,
         "on stations.STATION_ID = samples.STATION_ID) ",
         "left join NIVADATABASE.WATER_CHEMISTRY_VALUES chemval",
         "on samples.WATER_SAMPLE_ID = chemval.WATER_SAMPLE_ID " ,
         "where (stations.STATION_ID in (7579,7580,7581,7581) ",
         "and chemval.METHOD_ID in (1553,1554)) ",
         "group by stations.STATION_ID") %>% 
  paste(collapse = " ")
# edit(qry)
get_nivabase_data(qry)

# Test: insert station IDs into the query using sprintf 
qry_template <- c("select stations.STATION_ID, count(chemval.VALUE_ID) as count_chem",
         "from (NIVADATABASE.STATIONS stations left join NIVADATABASE.WATER_SAMPLES samples" ,
         "on stations.STATION_ID = samples.STATION_ID) ",
         "left join NIVADATABASE.WATER_CHEMISTRY_VALUES chemval",
         "on samples.WATER_SAMPLE_ID = chemval.WATER_SAMPLE_ID " ,
         "where (stations.STATION_ID in (%s) ",
         "and chemval.METHOD_ID in (1553,1554)) ",
         "group by stations.STATION_ID") %>% 
  paste(collapse = " ")

test_stations <- c(7579,7580,7581,7581)
qry <- sprintf(qry_template, paste(test_stations, collapse = ","))
get_nivabase_data(qry)
```

### c. Define query template
```{r}
# Using sprintf to insert both station IDs and method IDs into query
# The template has %s where ID strings will be inserted
qry_template <- c("select stations.STATION_ID, count(chemval.VALUE_ID) as count_temperature",
         "from (NIVADATABASE.STATIONS stations left join NIVADATABASE.WATER_SAMPLES samples" ,
         "on stations.STATION_ID = samples.STATION_ID)",
         "left join NIVADATABASE.WATER_CHEMISTRY_VALUES chemval",
         "on samples.WATER_SAMPLE_ID = chemval.WATER_SAMPLE_ID" ,
         "where (stations.STATION_ID in (%s)",
         "and chemval.METHOD_ID in (%s)) ",
         "group by stations.STATION_ID") %>% 
  paste(collapse = " ")

method_string <- paste(df_method$METHOD_ID, collapse = ",")

# Test: first 100 stations, all temperature methods
test_stations <- df_stations$STATION_ID[1:100]
qry <- sprintf(qry_template, 
               paste(test_stations, collapse = ","),
               method_string)
get_nivabase_data(qry)
```

### d. Do the actual counting   
Less than 1 minute  
```{r}
source("01fun_Explore_temp_data_functions.R")  # create_sequence_list()

# Make list (5 items) of STATION_ID strings
station_id_list <- create_sequence_list(df_stations$STATION_ID)

# Make list (5 items) of queries
qry_list <- station_id_list %>% map(~sprintf(qry_template, ., method_string))

# Make list (5 items) of results 
result_list <- qry_list %>% map(get_nivabase_data)

# Combine the 5 results in the list into a single result
result <- result_list %>% bind_rows()
nrow(result)
```

## 3. Save result
```{r}
openxlsx::write.xlsx(result, "Count of temperature stations in rivers.xlsx")
```

